---
title: "Mapping MeDIP counts matrix to Glioma SOLID patients"
author: ""
date: "April 14th, 2026"
output:
  pdf_document: default
  html_document: default
---

## Set up
```{r}
library(data.table) |> suppressMessages()
dir <- "/Users/julianguyen/Documents/SOLID-methylation/"
```

## Load in data

### MeDIP log2CPM matrix
```{r}
medip <- readRDS(paste0(dir, "data/rawdata/MeDIP_Solid_all_log2CPM.rds"))
print(dim(medip))
```

```{r}
medip[1:5,1:5]
```


```{r}
available_medip <- colnames(medip)
available_medip
```
These are the column names of the MeDIP counts matrix for all SOLID patients. There were 174 samples.


### Cholangio MeDIP log2CPM matrix
```{r}
cmat <- readRDS(paste0(dir, "data/rawdata/cholangio/MeDIP_Solid_cholangiocarcinoma_log2CPM.rds"))
print(dim(cmat))
```

```{r}
cholangio_medip <- colnames(cmat)
cholangio_medip
```

These should be all the cholangio patients. There are 56 samples total. This leaves 118 (174-56) MeDIP samples from the earlier matrix to be from the glioma patients.

### Identify Glioma MeDIP IDs

Removing the cholangio MeDIP sample IDs, these are what we are left with for glioma patients.

```{r}
glioma_medip <- available_medip[-which(available_medip %in% cholangio_medip)]
glioma_medip
```

## Subet for matched glioma SOLID IDs

Here are some metadata files I have available.

**Cholangio pheno**

```{r}
pheno <- fread(paste0(dir, "data/rawdata/cholangio/Pheno_cholangio_all_IDHwt_Vs_IDHmut.txt"))
dim(pheno) # 64 samples here
```

```{r}
head(pheno)
```

The MeDIP_ID column matches the column names, but only the cholangio matrices. I don't have an equivalent metadata file for the glioma samples.

```{r}
table(pheno$MeDIP_ID %in% available_medip)
```

```{r}
table(pheno$MeDIP_ID %in% glioma_medip)
```

**Glioma pheno**

Below is the metadata file from the tissue methylation which has the SOLID ids of the 29 glioma solid patients with methylation. I created a new column `mapping` which can be used to map to the MeDIP IDs.

```{r}
load(paste0(dir, "data/procdata/RGSets_targets.RData")) |> suppressMessages()
targets$mapping <- gsub("_0", "_", gsub("_0", "_", paste0("Solid_", sub(".*-", "", targets$Subject))))
head(targets)
```

### Method 1: subset for sampleIDs with 'SOLID' explicitly in sampleID

```{r}
glioma_with_solid <- glioma_medip[grep("Solid", glioma_medip)]
glioma_with_solid
```

```{r}
# keep only samples with '-1' (to filter for unique and presumably baseline samples)
glioma_with_solid[grepl("-1$", glioma_with_solid)]
```

Clean list of 10 available samples:

```{r}
clean_names <- sub("-.*", "", sub("batch6_", "", glioma_with_solid[grepl("-1$", glioma_with_solid)]))
print(clean_names)
print(clean_names %in% targets$mapping) # these are all in the list of glioma SOLID patients with methylation
```


### Method 2: assuming number following batch corresponds to SOLID patient ID (not sure if this is true)

```{r}
glioma_medip_rename <- sub("-.*", "", paste0("Solid_", sub("batch.*_", "", glioma_medip))) |> unique()
glioma_medip_rename
```

```{r}
table(glioma_medip_rename %in% targets$mapping) 
# 13 of these renamed sampleIDs are in the list of glioma SOLID patients with methylation
```

```{r}
glioma_medip_rename[glioma_medip_rename %in% targets$mapping]
```


